Investigating the Impact of Macroeconomic Indicators on the FTSE 250 Index Using Statistical \& Machine Learning Methods

### 2024-08-24 tex/paper.tex
introduction updated

on the FTSE 250 Index Using Machine Learning Methods}
\author{Mateen Khan}
\date{\today}

@@

The stock market has played a central role in the economic and 
political ascent of many nations ever since its inception (\citeboth{levine2002}).  
However, whenever stock market instabilities have arisen in history they 
have had far-reaching destabilising effects. The Mississippi Bubble was 
an early example of the inherent dangers within stock markets in the form of 
speculative bubbles (\citeboth{mississippibubble}). The role of 
macroeconomic policy in this crash was central, and the effects of the crash 
devastated the economy through rampant inflation. Nonetheless, stock markets also provide essential benefits by allowing companies to 
raise capital, offering investment opportunities for individuals, and 
supporting economic growth through efficient capital allocation (\citeboth{shahbaz2016}). Effective, sustainable
macroeconomic policy can greatly aid in this beneficial process. Therefore, it is essential 
for economic policymakers to understand the impact of macroeconomic changes 
on the stock market.

The sensitivity of the stock market to macroeconomic indicators is also of great 
relevance to investors. Changes in macroeconomic indicators have an effect on both consumer preferences
and business costs which impacts the profitability of firms. This 
leads to discrepancies between the valuation of firms by investors and 
stock prices, which in turn drives trends in the stock 
market price.

The significance of this study is to quantify the effects of macroeconomic 
indicators on the stock market as well as assessing the use of these indicators 
to forecast stock market prices over the long-term and short-term, both useful to 
economic policy makers and stock market investors respectively. 
Modelling of this kind typically involves choosing 
the most appropriate method from a variety: statistical analysis, 
econometric modelling, scenario planning, and machine learning 
techniques. Also, careful economic analysis is required to select
the macroeconomic indicators most relevant in 
the dynamics of a nation's stock market.

The objective of this study is to address how specific 
macroeconomic indicators influence stock market prices in a developed economy, and whether machine 
learning models can substitute traditional econometric models in predicting stock market prices. 
Ultimately, it aims to investigate two aspects of the influence of macroeconomic 
indicators on stock markets: 
\begin{enumerate}
    \item To examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250 
    \item To investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE250 when using a small, coarse-grain dataset.

\end{enumerate}

@@

The remainder of the paper is organised as follows. 
Section 2 presents a brief review of the literature in this problem space, 
and explains the rationale behind the selection of macroeconomic variables. 
Section 3 covers the data utilised in this study. Section 4 outlines the 
ARDL-ECM model and the results achieved. 
Section 5 constructs the RBFNN 
model, and presents the forecasting results. Section 6 concludes the study.

\section{Literature Review}

>>>>

on the FTSE 250 Index Using Statistical \& Machine Learning Methods}
\author{Mateen Khan}
\date{\today}

@@

The stock market is a significant component of the financial system and 
plays a crucial role in the broader economy. Consequently, economic policy 
makers take the stock market into account when making fiscal policy 
decisions using a combination of quantitative and qualitative methods. 
These methods such as econometric modelling and forecasting, and empirical 
analysis allow them to understand the short-term and long-term effects of 
macroeconomic indicators such as taxation rates, interest rates and exchange rates,
on the wider economy through the stock market.

Macroeconomic indicators can be effectively used in machine learning 
methods to forecast stock prices by serving as input features that capture 
the economic environment’s influence on market behaviour. By carefully 
selecting the most relevant indicators and preprocessing data, 
as well as choosing appropriate models, machine learning can enhance the accuracy of stock price forecasts (\citeboth{Sarma2023,Prasad2023}).

The significance of this study is to quantify the effects of macroeconomic indicators on the stock market as well as assessing the use of these indicators to forecast stock market prices over the short-term and long-term. In addition, the study seeks to understand which macroeconomic indicators are most relevant in capturing the dynamics of a stock market to aid forecasting.

In particular, the objectives of this study are to address how specific macroeconomic indicators influence stock market prices in a developed economy, and whether machine learning models can substitute traditional econometric models in predicting stock market prices. It aims to:
\begin{enumerate}
    \item examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
    \item investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE250 when using a small, coarse-grain dataset.
\end{enumerate}

@@

The remainder of this paper will be organised as follows. 
Section 2 provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
Section 3 presents the methodology of the paper. Section 4 discusses the results and Section 5 concludes
the study.

\section{Literature Review}
\label{sec:lit}
